Learning to Pivot with Adversarial Networks

### 2017-02-01 tex/icml2017/paper.tex
save

Many inference problems involve data generation processes that are not uniquely
specified or are uncertain in some way. In a scientific context, the presence of
several plausible data generation processes is often associated to the presence
of systematic uncertainties. Robust inference is possible if it is based on a
pivot -- a quantity whose distribution is invariant to the unknown value of the
(categorical or continuous) nuisance parameters that parametrizes this family of
generation processes. In this work,  we introduce a flexible training procedure
based on adversarial networks for enforcing the pivotal property on a predictive
model. We derive theoretical results showing that the proposed algorithm tends towards a minimax
solution corresponding to a predictive model that is both optimal and
independent of the nuisance parameters (if that models exists) or for which one
can tune the trade-off between power and robustness. Finally, we demonstrate the
effectiveness of this approach with a toy example and an example from particle
physics.

@@

    \begin{tikzpicture}[shorten >= 1pt, ->, node distance=\layersep,scale=.85]
    \tikzstyle{neuron} = [circle, minimum size=0.25cm, draw=black!30, line width=0.3mm, fill=white]

@@

    \foreach \source in {1,...,3}
        \foreach \dest in {1,...,5}
            \path[black!20] (f-I-\source) edge (f-H1-\dest);

    \foreach \source in {1,...,5}
        \path[black!30] (f-H2-\source) edge (f-O);

    \node[black!30] at (1.5,-3) {...};

@@

    \foreach \source in {1,...,1}
        \foreach \dest in {1,...,5}
            \path[black!20] (r-I-\source) edge (r-H1-\dest);

    \foreach \source in {1,...,5}
        \path[black!30] (r-H2-\source) edge (r-O1);
    \foreach \source in {1,...,5}
        \path[black!30] (r-H2-\source) edge (r-O2);
    \foreach \source in {1,...,5}
        \path[black!30] (r-H2-\source) edge (r-O3);

    \node[black!30] at (9,-3) {...};

>>>>

Many inference problems involve data generation processes that are not uniquely
specified or are uncertain in some way. In a scientific context, the presence of
several plausible data generation processes is often associated to the presence
of systematic uncertainties. Robust inference is possible if it is based on a
pivot -- a quantity whose distribution is invariant to the unknown value of the
(categorical or continuous) nuisance parameters that parametrizes this family of
generation processes. In this work,  we introduce a flexible training procedure
based on adversarial networks for enforcing the pivotal property on a predictive
model. We derive theoretical results showing that the proposed algorithm tends
towards a minimax solution corresponding to a predictive model that is both
optimal and independent of the nuisance parameters (if that model exists) or
for which one can tune the trade-off between power and robustness. Finally, we
demonstrate the effectiveness of this approach with a toy example and an example
from particle physics.

@@

    \begin{tikzpicture}[shorten >= 1pt, ->, node distance=\layersep,scale=.85]
    \tikzstyle{neuron} = [circle, minimum size=0.25cm, draw=black!20, line width=0.3mm, fill=white]

@@

    \foreach \source in {1,...,3}
        \foreach \dest in {1,...,5}
            \path[black] (f-I-\source) edge (f-H1-\dest);

    \foreach \source in {1,...,5}
        \path[black] (f-H2-\source) edge (f-O);

    \node[black] at (1.5,-3) {...};

@@

    \foreach \source in {1,...,1}
        \foreach \dest in {1,...,5}
            \path[black] (r-I-\source) edge (r-H1-\dest);

    \foreach \source in {1,...,5}
        \path[black] (r-H2-\source) edge (r-O1);
    \foreach \source in {1,...,5}
        \path[black] (r-H2-\source) edge (r-O2);
    \foreach \source in {1,...,5}
        \path[black] (r-H2-\source) edge (r-O3);

    \node[black] at (9,-3) {...};
